### 2014-11-10 RelatedWork.tex
stuff
Pass through RelatedWork

\section{Related Work}
\label{relatedWork}

Hardware verification is dominated by model-checking; for instance, processor
verification ~\cite{burch1994automatic, mcmillan1998verification} and more
recently, Intel's execution cluster verification~\cite{kaivola2009replacing}.
Many abstraction techniques are used to reduce designs to finite state spaces,
which can be explored exhaustively. There are limits to the construction of
sound abstractions, so verifications of, e.g., cache-coherence protocols have
almost always treated systems with concrete topologies, involving particular
finite numbers of caches and processors. For instance, explicit-state model
checking tools like Murphi~\cite{murphi} or TLC~\cite{tlc} are only able to
handle small and simplified systems, \eg{} a single-level cache hierarchy with
three CPUs, two addresses, and one bit of data. Symbolic model checking
techniques have fared better: McMillan \etal{} have verified a 2-level MSI
protocol based on the Gigamax distributed multiprocessor using
SMV~\cite{gigamax}. Optimizations on these techniques (\eg{} partial order
reduction~\cite{part}, symmetry reduction~\cite{sym1, sym2}, compositional
reasoning~\cite{comp, Mccomp, mcc}, \etc{}) scale the approach further, but
they are still unable to handle realistic hierarchical protocols. We might hope
to achieve higher assurance and understanding of design ideas by verifying
\emph{infinite families} of hardware designs, which resist reduction to
finite-state models.  A potential related benefit is reducing the computational
resources required to formally verify a new design in the family, applying a
more general theorem without much new analysis.

Theorem provers have also been used to verify microprocessors, \eg{} HOL has
been used to verify an academic microprocessor AVI-1 \cite{windley1995formal}.
Term-rewriting based techniques have been used to specify microprocessors, with
the goal of making them more amenable to verification \cite{shen1999using}.

Theorem provers have also been used in verifying cache coherence protocols. The
correctness of the FLASH cache coherence protocol~\cite{flash} for a
single-level hierarchy was proven with PVS~\cite{park} using a mapping, called
``aggregation of transactions,'' from multiple localized transactions to a
single global transaction, and finally using model checking to prove the
correctness of the global transactions model. However, they do not provide a
methodology to prove the aggregation of transactions for other protocols, which
is where, we believe, the difficulty lies; in contrast, our modular proof
relies on generic invariants, making it more adaptable to different protocols.
Cachet~\cite{StoyShenArvind:Proofs}, another single-level cache coherence
protocol, was verified to be obey the CRF memory model~\cite{Shen:CRF}, and
parts of the proof were implemented using PVS.

@@

There has been a lot of work in providing frameworks for axiomatic
specifications of memory models based on empirically observed behavior of
certain ``litmus
tests''\cite{Alglave11,Alglave:FMSD,alglave2012formal,Alglave:2010,Alglave:TACAS,alglave2014herding}.
There has been some work in providing operational and axiomatic semantics that
describes the observed behavior (based on certain ``litmus tests'' of real
hardware
\cite{mador2012axiomatic,sarkar2011understanding,sarkar2012synchronising}.
Finally, there has been work formally relating the axiomatic specification of
TSO to operational semantics \cite{x86tsocacm10}.

>>>>

\vspace{-.4cm}
\section{Related Work}
\label{relatedWork}

Hardware verification is dominated by model checking; for instance, processor
verification ~\cite{burch1994automatic, mcmillan1998verification} and more
recently, Intel's execution cluster verification~\cite{kaivola2009replacing}.
Many abstraction techniques are used to reduce designs to finite state spaces,
which can be explored exhaustively. There are limits to the construction of
sound abstractions, so verifications of, e.g., cache-coherence protocols have
almost always treated systems with concrete topologies, involving particular
finite numbers of caches and processors. For instance, explicit-state model
checking tools like Murphi~\cite{murphi} or TLC~\cite{tlc} are only able to
handle small and simplified systems, \eg{} a single-level cache hierarchy with
three CPUs, two addresses, and one bit of data. Symbolic model-checking
techniques have fared better: McMillan \etal{} have verified a 2-level MSI
protocol based on the Gigamax distributed multiprocessor using
SMV~\cite{gigamax}. Optimizations on these techniques (\eg{} partial order
reduction~\cite{part}, symmetry reduction~\cite{sym1, sym2}, compositional
reasoning~\cite{comp, Mccomp, mcc}, \etc{}) scale the approach further, but
they are still unable to handle realistic hierarchical protocols. We might hope
to achieve higher assurance and understanding of design ideas by verifying
\emph{infinite families} of hardware designs, which resist reduction to
finite-state models.  A potential related benefit is reducing the computational
resources required to formally verify a new design in the family, applying a
more general theorem without much new analysis.

Theorem provers have also been used to verify microprocessors, \eg{} HOL has
been used to verify an academic microprocessor AVI-1 \cite{windley1995formal}.
Term-rewriting techniques have been used to specify microprocessors, with
the goal of making them more amenable to verification \cite{shen1999using}.

Cache-coherence proofs have also been done with mechanized theorem provers. The
correctness of the FLASH cache-coherence protocol~\cite{flash} for a
single-level hierarchy was proven with PVS~\cite{park} using a mapping, called
``aggregation of transactions,'' from multiple localized transactions to a
single global transaction, and finally using model checking to prove the
correctness of the global transactions model. However, they do not provide a
methodology to prove the aggregation of transactions for other protocols, which
is where, we believe, the difficulty lies; in contrast, our modular proof
relies on generic invariants, making it more adaptable to different protocols.
Cachet~\cite{StoyShenArvind:Proofs}, another single-level cache coherence
protocol, was verified to be obey the CRF memory model~\cite{Shen:CRF}, and
parts of the proof were implemented using PVS.

@@

Before verifying that a hardware design implements an ISA-level semantics, it
is necessary to decide what the correct semantics is.  Much recent work has
studied widely deployed hardware platforms empirically to derive formal
semantics.  While our case study in this paper hews to the uncontroversial
sequential-consistency model, the empirical challenges arise in characterizing
weaker memory models.  Approaches have included building axiomatic
memory-model specifications based on empirically observed behavior on
certain ``litmus
tests''~\cite{Alglave11,Alglave:FMSD,alglave2012formal,Alglave:2010,Alglave:TACAS,alglave2014herding},
building models that include operational semantics~\cite{mador2012axiomatic,sarkar2011understanding,sarkar2012synchronising}, and relating the axiomatic specification of
TSO to operational semantics~\cite{x86tsocacm10}.
